---
title: "DA4 -  Assignment 1"
author: "Tamas Koncz"
date: '2018-02-11'
output:
  html_document:
    df_print: paged
  html_notebook:
    df_print: paged
---

```{r setup, message=FALSE, include=FALSE}
require(data.table)
require(ggplot2)
require(gridExtra)
require(caret)

options(scipen = 999)

theme_set(theme_minimal())   # globally set ggplot theme

set.seed(1234)
RMSE <- function(x, true_x) sqrt(mean((x - true_x)^2))
ERR_SQ <- function(x, true_x) (x - true_x)^2
```


```{r, include= FALSE}
data <- fread("airbnb_london_workfile.csv",
              stringsAsFactors = FALSE)
```
 
 
 
 
  
  
#### Code snippet for selecting random borough 

Below code was used to pick a random borough from the dataset.
Here, and in later parts as well, where a random generator is involved in producing the results, I used set.seed(93) to help foster reproducibility of results.
```{r}
boroughs <- data[, .(count = .N, 
                     avg_price = mean(price)), 
                 keyby = f_neighbourhood_cleansed][order(-count)]

boroughs[, borough := factor(f_neighbourhood_cleansed, 
                             levels = boroughs[order(count)][, f_neighbourhood_cleansed])]
boroughs[, f_neighbourhood_cleansed := NULL]

# randomly picking an area > 1000
set.seed(93) #for reproducibility
selected <- sample(boroughs[count > 1000]$borough, 1)

```


```{r, fig.align= 'center', fig.width= 10, echo= FALSE}
max_count <- boroughs[, max(count)]
max_avg_price <- boroughs[, max(avg_price)]

boroughs[borough == selected, ] #TODO: make this bold on the chart

ggplot(data = boroughs) + 
  geom_bar(data = boroughs[borough == selected], aes(x = borough, y = 6000), fill = "lightblue",  stat = "identity") +
  geom_point(aes(x = borough, y = count, color = "# of Observations"), shape = 20, size = 2) +
  geom_segment(aes(x = borough, y = count, xend = borough, yend =0, color = "# of Observations")) + 
  geom_point(aes(x = borough, y = avg_price * (max_count/max_avg_price), color = "Avg. Price"), shape = 4, size = 2) + 
  scale_y_continuous(limits = c(0,6000), sec.axis = sec_axis(~./(max_count/max_avg_price), name = "Avg. Price")) + 
  scale_color_manual(name = "Legend", values = c("# of Observations" = "tomato", "Avg. Price" = "darkblue")) + 
  guides(color=guide_legend(override.aes=list(shape=15))) +
  labs(y = "# of Observations", x = "Borough") +
  coord_flip()
```

```{r, include= FALSE}
london <- copy(data)
rm(data)
rm(boroughs)
rm(max_count)
rm(max_avg_price)
```

#### Data cleaning

The below steps were applied to prepare the data for analysis.

    1. Handling missing (NA) values: 
```{r, echo= FALSE}
#handling NAs
missing_values <- as.data.table(t(london[, lapply(.SD, function(x) sum(is.na(x))), .SDcols = names(london)]), keep.rownames=TRUE)
setnames(missing_values, c("variable", "NA.Count"))

missing_values[order(-NA.Count)][NA.Count>0]
```

I followed the logic of checking if there seems to be any significant relationship between variables with many NA-s and price (example graph below).
As I did not see anything meaningful, I decided to drop these variables. There might be a way to guess their values based on the data, however, given the lack of expected predicting power, I took the "clean" path.

```{r, echo= FALSE, fig.height=4, fig.align='right', fig.width=5}
ggplot(data= london, aes(x = n_review_scores_rating, y = price)) + 
  geom_point() + 
  geom_smooth(method='lm')
```

Upon checking airbnb.com, I decided that the missing cleaning fees _actually_ mean 0 added cost for this item, hence I replaced NAs with zero for this one variable:
```{r}
london[, cleaning_fee := ifelse(is.na(usd_cleaning_fee), 0, usd_cleaning_fee)]
```

And then just proceeded to drop the many-NA variables:
```{r}
london[, c("usd_cleaning_fee", 
           "n_review_scores_rating",
           "n_reviews_per_month",
           "n_days_since",
           "p_host_response_rate") := NULL]
```

We are almost done - for the reamining two variables with some NAs, I was comfortable dropping the whole observation row, as their numbers are very limited compared to the whole dataset:
```{r}
london <- london[complete.cases(london)]
```

    2. Getting rid of unnecessary variables

There were some variables which encoded duplicates, I just dropped all these values:
```{r}
london[, c("neighbourhood_cleansed", 
           "property_type", 
           "room_type", 
           "usd_price_day", 
           "cancellation_policy") := NULL]
```


```{r, include= FALSE}
factor_cols <- names(london)[names(london) %like% "^f_.*"]
london[, (factor_cols) := lapply(.SD, as.factor), .SDcols = factor_cols]
```

    3. Feature engineering
The last step of data preparation was creating a few set of additional variables that I expect to improve our models' fit.

The first one was to create log(...) transformed version of price - this will be explained in detail in the next section.
```{r}
london[, log_price:= log(price)]
```

I've created another variable for the dummies - this one however is a total "score" (summing up all available extras).  
This is a very simple approach, without any weightings on different benefits - however, it can provide some insight into how people value non-core offerings.
```{r}
london[, d_total := Reduce("+", .SD), .SDcols = names(london) %like% "^d_.*"]
```

For some variables, I've added second polinomial terms, for better fit. More explanation in the visual section to follow.
```{r}
# TODO: polynomial terms! 
```


```{r, include = FALSE}
rm(missing_values)
```

```{r, include = FALSE}
#creating subset
kensington_chelsea <- london[f_neighbourhood_cleansed == selected]
```

#### Data exploration

```{r, echo= FALSE}
ggplot(data= london, aes(x = log_price)) + geom_histogram()
ggplot(data= kensington_chelsea, aes(x = log_price)) + geom_histogram()

# TODO: add reasoning for log transformation. two hist, not-log, common density, log

ggplot(data= london, aes(x = n_accommodates)) + geom_histogram()
ggplot(data= london, aes(x = n_accommodates, y = log_price)) + geom_point() + geom_smooth()

ggplot(data= london, aes(x = n_beds)) + geom_histogram()
ggplot(data= london, aes(x = n_beds, y = log_price)) + geom_point() + geom_smooth()
#transform both for deminishing returns?

# f_room_type
ggplot(data= london, aes(x = f_room_type)) + geom_bar()
ggplot(data= london, aes(x = f_room_type, y = log_price)) + geom_boxplot()

# f_cancellation_policy -> surprising result. correlation with sth else?
ggplot(data= london, aes(x = f_cancellation_policy)) + geom_bar()
ggplot(data= london, aes(x = f_cancellation_policy, y = log_price)) + geom_boxplot()

#d_breakfast -> no big impact. add d_ to TOTAL?
ggplot(data= london, aes(x = d_breakfast)) + geom_bar()
ggplot(data= london, aes(x = factor(d_breakfast), y = log_price)) + geom_boxplot()

#n_number_of_reviews
ggplot(data= london, aes(x = n_number_of_reviews)) + geom_histogram()
ggplot(data= kensington_chelsea, aes(x = n_number_of_reviews, y = log_price)) + geom_point() + geom_smooth()

#n_minimum_nights --> some outliers, handle
ggplot(data= london, aes(x = n_minimum_nights)) + geom_histogram()
ggplot(data= london, aes(x = n_minimum_nights, y = log_price)) + geom_point() + geom_smooth()

#f_property_type --> some interaction with other variables, eg. room type?
ggplot(data= london, aes(x = f_property_type)) + geom_bar()
ggplot(data= london, aes(x = f_property_type, y = log_price)) + geom_boxplot()

# usd_cleaning_fee --> model non-linearity?
ggplot(data= london, aes(x = cleaning_fee)) + geom_histogram()
ggplot(data= london, aes(x = cleaning_fee, y = log_price)) + geom_point() + geom_smooth()


# d_total
ggplot(data= london, aes(x = d_total)) + geom_histogram()
ggplot(data= london, aes(x = d_total, y = log_price)) + geom_point() + geom_smooth()
```

#### Creating a hold-out test set

I'm create separate datasets for model training and performance evaluations (same method was followed for the Kensington and Chelsea subsample), using basic Caret functions.  
Model selection will be done via cross-validation on the training set.
```{r}
training_ratio <- 0.7

set.seed(93) #for reproducibility
train_indices <- createDataPartition(y = london[["log_price"]],
                                     times = 1,
                                     p = training_ratio,
                                     list = FALSE)
london_train <- london[train_indices, ]
london_test <- london[-train_indices, ]
```

```{r, include= FALSE}
set.seed(93) #for reproducibility
train_indices <- createDataPartition(y = kensington_chelsea[["log_price"]],
                                     times = 1,
                                     p = training_ratio,
                                     list = FALSE)
kensington_chelsea_train <- kensington_chelsea[train_indices, ]
kensington_chelsea_test <- kensington_chelsea[-train_indices, ]
```
 
 
#### Setting the control parameters for 10-fold CV  

```{r}
fit_control <- trainControl(method = "cv", number = 10)
```

#### Model training

I'm fitting four separte models - they share the same predictors and train(...) parameters. Hence, I'll only include code for the whole London dataset - what happens for the subsample is analogous.  
  
The first model is very basic - it just uses one predictor, n_accommodates. It is expected to be used rather as a benchmark, than an actual model for prediction.
```{r}
set.seed(93) #for reproducibility
model_1_london <- train(log_price ~ n_accommodates, 
                   data = london_train, 
                   method = "lm", 
                   trControl = fit_control)
```

```{r, include= FALSE}
set.seed(93) #for reproducibility
model_1_kensington_chelsea <- train(log_price ~ n_accommodates, 
                   data = kensington_chelsea_train, 
                   method = "lm", 
                   trControl = fit_control)
```


```{r}
#simple model - using multiple predictor variables

set.seed(93) #for reproducibility
model_2_london <- train(log_price ~ n_accommodates + n_beds + n_bathrooms + 
                          f_property_type + f_room_type + f_bed_type, 
                   data = london_train, 
                   method = "lm", 
                   trControl = fit_control)
```

```{r, include= FALSE}

set.seed(93) #for reproducibility
model_2_kensington_chelsea <- train(log_price ~ n_accommodates + n_beds + n_bathrooms +
                                            f_property_type + f_room_type + f_bed_type, 
                                          data = kensington_chelsea_train, 
                                          method = "lm",
                                          trControl = fit_control)
```

```{r}
#simple model - using multiple predictor variables

set.seed(93) #for reproducibility
model_3_london <- train(log_price ~ n_accommodates + n_beds + n_bathrooms + 
                          f_property_type + f_room_type + f_bed_type, 
                   data = london_train, 
                   method = "lm", 
                   trControl = fit_control)
```

```{r, include= FALSE}

set.seed(93) #for reproducibility
model_3_kensington_chelsea <- train(log_price ~ n_accommodates + n_beds + n_bathrooms +
                                            f_property_type + f_room_type + f_bed_type, 
                                          data = kensington_chelsea_train, 
                                          method = "lm",
                                          trControl = fit_control)
```


Setting up LASSO, and hyperparameter-tuning with Caret:  
```{r}
tune_grid <- expand.grid("alpha" = 1,
                             "lambda" = seq(0, 0.015, 0.001))
```


```{r, include= FALSE}
l <- sapply(kensington_chelsea_train, function(x) is.factor(x))
m <- kensington_chelsea_train[, ..l]

sapply(kensington_chelsea_train, function(x) length(unique(x)))

rm(l)
rm(m)
```


```{r}
#full model - most complexity, all variables

set.seed(93) #for reproducibility
model_4_london <- train(log_price ~ . -price -d_total -f_neighbourhood_cleansed -d_washerdryer -d_freeparkingonstreet -d_paidparkingoffpremises, 
                    data = london_train, 
                    method = "glmnet",
                    preProcess = c("center", "scale"),
                    tuneGrid = tune_grid,
                    metric = "RMSE",
                    trControl = fit_control)
```


```{r}
#full model - most complexity, all variables

set.seed(93) #for reproducibility
model_4_kensington_chelsea <- train(log_price ~ . -price -d_total -f_neighbourhood_cleansed -d_washerdryer -d_freeparkingonstreet -d_paidparkingoffpremises, 
                    data = kensington_chelsea_train, 
                    method = "glmnet",
                    preProcess = c("center", "scale"),
                    tuneGrid = tune_grid,
                    metric = "RMSE",
                    trControl = fit_control)
```

#### Selecting the best model

RMSE values from the different models are below:
```{r, echo= FALSE, fig.align='center', fig.width=10, fig.height=5}
model_1_london_rmse_cv <- model_1_london$results[["RMSE"]]
model_2_london_rmse_cv <- model_2_london$results[["RMSE"]]
model_3_london_rmse_cv <- model_3_london$results[["RMSE"]]
model_4_london_rmse_cv <- min(model_4_london$results[["RMSE"]])

model_1_kensington_chelsea_rmse_cv <- model_1_kensington_chelsea$results[["RMSE"]]
model_2_kensington_chelsea_rmse_cv <- model_2_kensington_chelsea$results[["RMSE"]]
model_3_kensington_chelsea_rmse_cv <- model_3_kensington_chelsea$results[["RMSE"]]
model_4_kensington_chelsea_rmse_cv <- min(model_4_kensington_chelsea$results[["RMSE"]])

m_logrmse <- matrix(c(1, model_1_london_rmse_cv, model_1_kensington_chelsea_rmse_cv,
                   2, model_2_london_rmse_cv, model_2_kensington_chelsea_rmse_cv,
                   3, model_3_london_rmse_cv, model_3_kensington_chelsea_rmse_cv,
                   4, model_4_london_rmse_cv, model_4_kensington_chelsea_rmse_cv), 
                 nrow= 4, byrow= TRUE)
colnames(m_logrmse) <- c('model', 'London RMSE', "Kensington and Chelsea RMSE")
logrmse.table <- data.table(m_logrmse)
ggplot(data= logrmse.table) +
  geom_point(aes(x= model, y= as.numeric(`London RMSE`), color= "London"), stat="identity") +
  geom_line(aes(x= model, y= as.numeric(`London RMSE`), color= "London"), stat="identity") +
  geom_point(aes(x= model, y= as.numeric(`Kensington and Chelsea RMSE`), color= "Kensington and Chelsea"), stat="identity") +
  geom_line(aes(x= model, y= as.numeric(`Kensington and Chelsea RMSE`), color= "Kensington and Chelsea"), stat="identity") +
  labs(y = "RMSE of log predictions", x = "Model", title= "RMSE values of log-prediction, after cross-validation, training sample") +
  theme(legend.position=c(1, 1), 
        legend.title=element_blank(),
        legend.background = element_rect(fill="white", size=0.5, linetype="solid"))
```
However, these are RMSE-s of the log_price predictions - while we are after predicting the actual prices. Let's fix this.  

As the first step, we'll make predictions on the trainings set with all our models. Reference for how to calculate these predictions predictions can be found here: https://www.r-bloggers.com/forecasting-from-log-linear-regressions/.  
  
  
Implementations is via the below code, which calculated the log, uncorrected level, and corrected level predictions for a given model on a given dataset:
```{r}
predictions_calc <- function(model, dt) {
  log_pred <- predict(model, dt)
  n <- dt[,.N]
  k <- length(predictors(model))
  s2 <- sum((log_pred - dt$log_price)^2) / (n-k)
  
  log_pred <- data.table(log_pred)
  pred_uncorrected <- data.table(exp(log_pred))
  pred <- data.table(exp(log_pred + s2 / 2))
  
  dt_pred <- cbind(log_pred, pred_uncorrected, pred)
  setnames(dt_pred, c("Log Prediction", "Uncorrected Prediction", "Prediction"))
  
  return(dt_pred)
}
```


Calculating RMSE based on the level predictions (both the uncorrected and the corrected):
```{r}
model_1_london_pred <- predictions_calc(model_1_london, london_train)

model_1_london_rmse_uncorr_train <- RMSE(model_1_london_pred$`Uncorrected Prediction`, london_train$price)
model_1_london_rmse_train <- RMSE(model_1_london_pred$`Prediction`, london_train$price)
```

To save ink-space, the coden for the other 7 model is not shown, but the exact same logic is applied to those as well to calculate RMSEs.

```{r echo= FALSE}
#1
model_1_kensington_chelsea_pred <- predictions_calc(model_1_kensington_chelsea, kensington_chelsea_train)
model_1_kensington_chelsea_rmse_uncorr_train <- RMSE(model_1_kensington_chelsea_pred$`Uncorrected Prediction`, kensington_chelsea_train$price)
model_1_kensington_chelsea_rmse_train <- RMSE(model_1_kensington_chelsea_pred$`Prediction`, kensington_chelsea_train$price)

#2
model_2_london_pred <- predictions_calc(model_2_london, london_train)
model_2_london_rmse_uncorr_train <- RMSE(model_2_london_pred$`Uncorrected Prediction`, london_train$price)
model_2_london_rmse_train <- RMSE(model_2_london_pred$`Prediction`, london_train$price)

model_2_kensington_chelsea_pred <- predictions_calc(model_2_kensington_chelsea, kensington_chelsea_train)
model_2_kensington_chelsea_rmse_uncorr_train <- RMSE(model_2_kensington_chelsea_pred$`Uncorrected Prediction`, kensington_chelsea_train$price)
model_2_kensington_chelsea_rmse_train <- RMSE(model_2_kensington_chelsea_pred$`Prediction`, kensington_chelsea_train$price)

#3
model_3_london_pred <- predictions_calc(model_3_london, london_train)
model_3_london_rmse_uncorr_train <- RMSE(model_3_london_pred$`Uncorrected Prediction`, london_train$price)
model_3_london_rmse_train <- RMSE(model_3_london_pred$`Prediction`, london_train$price)

model_3_kensington_chelsea_pred <- predictions_calc(model_3_kensington_chelsea, kensington_chelsea_train)
model_3_kensington_chelsea_rmse_uncorr_train <- RMSE(model_3_kensington_chelsea_pred$`Uncorrected Prediction`, kensington_chelsea_train$price)
model_3_kensington_chelsea_rmse_train <- RMSE(model_3_kensington_chelsea_pred$`Prediction`, kensington_chelsea_train$price)

#4
model_4_london_pred <- predictions_calc(model_4_london, london_train)
model_4_london_rmse_uncorr_train <- RMSE(model_4_london_pred$`Uncorrected Prediction`, london_train$price)
model_4_london_rmse_train <- RMSE(model_4_london_pred$`Prediction`, london_train$price)

model_4_kensington_chelsea_pred <- predictions_calc(model_4_kensington_chelsea, kensington_chelsea_train)
model_4_kensington_chelsea_rmse_uncorr_train <- RMSE(model_4_kensington_chelsea_pred$`Uncorrected Prediction`, kensington_chelsea_train$price)
model_4_kensington_chelsea_rmse_train <- RMSE(model_4_kensington_chelsea_pred$`Prediction`, kensington_chelsea_train$price)
```


Let's visualize the results to facilitate our selection of the best model:
```{r, echo=FALSE, fig.width= 15}
m_rmse_uncorr <- matrix(c(1, model_1_london_rmse_uncorr_train, model_1_kensington_chelsea_rmse_uncorr_train, 
                   2, model_2_london_rmse_uncorr_train, model_2_kensington_chelsea_rmse_uncorr_train, 
                   3, model_3_london_rmse_uncorr_train, model_3_kensington_chelsea_rmse_uncorr_train, 
                   4, model_4_london_rmse_uncorr_train, model_4_kensington_chelsea_rmse_uncorr_train), 
                 nrow= 4, byrow= TRUE)


colnames(m_rmse_uncorr) <- c('model', 'London RMSE', "Kensington and Chelsea RMSE")
m_rmse_uncorr.table <- data.table(m_rmse_uncorr)
p1 <- ggplot(data= m_rmse_uncorr.table) +
  geom_point(aes(x= model, y= as.numeric(`London RMSE`), color= "London"), stat="identity", size= 3) +
  geom_line(aes(x= model, y= as.numeric(`London RMSE`), color= "London"), stat="identity", size= 1.5) +
  geom_point(aes(x= model, y= as.numeric(`Kensington and Chelsea RMSE`), color= "Kensington and Chelsea"), stat="identity", size= 3) +
  geom_line(aes(x= model, y= as.numeric(`Kensington and Chelsea RMSE`), color= "Kensington and Chelsea"), stat="identity", size= 1.5) +
  labs(y = "RMSE of uncorrected predictions", x = "Model", title= "RMSE values of uncorrected level-prediction,\nafter cross-validation, training sample") +
  theme(legend.position="none")

m_rmse <- matrix(c(1, model_1_london_rmse_train, model_1_kensington_chelsea_rmse_train, 
                   2, model_2_london_rmse_train, model_2_kensington_chelsea_rmse_train, 
                   3, model_3_london_rmse_train, model_3_kensington_chelsea_rmse_train, 
                   4, model_4_london_rmse_train, model_4_kensington_chelsea_rmse_train), 
                 nrow= 4, byrow= TRUE)


colnames(m_rmse) <- c('model', 'London RMSE', "Kensington and Chelsea RMSE")
m_rmse.table <- data.table(m_rmse)
p2 <- ggplot(data= m_rmse.table) +
  geom_point(aes(x= model, y= as.numeric(`London RMSE`), color= "London"), stat="identity", size= 3) +
  geom_line(aes(x= model, y= as.numeric(`London RMSE`), color= "London"), stat="identity", size= 1.5) +
  geom_label(aes(x= model, y= as.numeric(`London RMSE`), label= round(as.numeric(`London RMSE`),2)), hjust= 0, vjust= 0) +
  geom_point(aes(x= model, y= as.numeric(`Kensington and Chelsea RMSE`), color= "Kensington and Chelsea"), stat="identity", size= 3) +
  geom_line(aes(x= model, y= as.numeric(`Kensington and Chelsea RMSE`), color= "Kensington and Chelsea"), stat="identity", size= 1.5) +
  geom_label(aes(x= model, y= as.numeric(`Kensington and Chelsea RMSE`), label= round(as.numeric(`Kensington and Chelsea RMSE`),2)), hjust= 0, vjust= 0) +
  labs(y = "RMSE of predictions", x = "Model", title= "RMSE values of level-prediction,\nafter cross-validation, training sample") +
  theme(legend.position=c(1, 1), 
        legend.title=element_blank(),
        legend.background = element_rect(fill="white", size=0.5, linetype="solid"))

grid.arrange(p1, p2, ncol = 2)
```



#### Evaluating the best model
After evaluating the different models based on different criteria, and selecting the best ones, let's see how they perform on the held-out test set:
```{r}
#TODO
```


#### London prices 2018

#### Budapest prices 2018
